Simulating High-Dimensional Multivariate Data

### 2020-12-29 020-simulation-algorithm.Rmd

1. Mapping step  
   i. Convert the target Spearman correlation matrix ${\bf R_{Spearman}}$ to the corresponding MVN Pearson correlation ${\bf R_X}$.   
   or  
i$^\prime$. Convert the target Kendall $\tau$ matrix ${\bf R_{Kendall}}$ to the corresponding MVN Pearson correlation ${\bf R_X}$.
2. Checking admissibility and nearest correlation computation
    (i) Check that ${\bf R_{X}}$ is a correlation matrix, a PSD matrix with 1's along the diagonal.   
	(ii) If ${\bf R_{X}}$ is a correlation matrix, the input matrix is *admissible* in this scheme.  
	Otherwise,  
(ii$^\prime$) Replace ${\bf R_{X}}$ with the nearest correlation matrix ${\bf \widetilde{R}_{X}}$, in the Frobenius norm.
3. NORTA step
	(i) Generate ${\bf X}=(X_1, \ldots, X_d) \sim N_d({\bf 0}, {\bf R_{X}})$.  
	(ii) Transform ${\bf X}$ to ${\bf U} = (U_1, \ldots,  U_d)$ viz $U_i=\Phi(X_i)$, $i=1, \ldots, d$.  
	(iii) Return ${\bf Y}  = (Y_1, \ldots,  Y_d)$, where $Y_i=F_i^{-1}(U_i)$, $i=1, \ldots, d$;

@@

Yet this scheme still suffers similar issues as the original Pearson-matching NORTA algorithm regarding the increasing probability of encountering a non-admissible correlation matrix as dimension increases.
In our experience, the *mapping step* for large $d$ almost always produces a $R_X$ that is not in $\Omega_P$ (or even a correlation matrix). The second step handles this issue by computing the nearest correlation matrix to $R_X$.
In the RNA-seq example (Section \@ref(examples)), we demonstrate this and show how to quantify and control the approximation error.

>>>>

1. Mapping step  
   i. Convert the target Spearman correlation matrix $R_S$ to the corresponding MVN Pearson correlation $R_X$.   
   or  
i$^\prime$. Convert the target Kendall $\tau$ matrix $R_K$ to the corresponding MVN Pearson correlation $R_X$.
2. Checking admissibility and nearest correlation computation
    (i) Check that $R_X$ is a correlation matrix, a PSD matrix with 1's along the diagonal.   
	(ii) If $R_X$ is a correlation matrix, the input matrix is *admissible* in this scheme.  
	Otherwise,  
(ii$^\prime$) Replace $R_X$ with the nearest correlation matrix $R_X$, in the Frobenius norm.
3. NORTA step
	(i) Generate ${\bf X}=(X_1, \ldots, X_d) \sim N_d({\bf 0}, R_X)$.  
	(ii) Transform ${\bf X}$ to ${\bf U} = (U_1, \ldots,  U_d)$ viz $U_i=\Phi(X_i)$, $i=1, \ldots, d$.  
	(iii) Return ${\bf Y}  = (Y_1, \ldots,  Y_d)$, where $Y_i=F_i^{-1}(U_i)$, $i=1, \ldots, d$;

@@

Yet this scheme still suffers similar issues as the original Pearson-matching NORTA algorithm regarding the increasing probability of encountering a non-admissible correlation matrix as dimension increases.
In our experience, the *mapping step* for large $d$ almost always produces a $R_X$ that is not in $\Omega_P$ (or even a correlation matrix). The second step handles this issue by computing the nearest correlation matrix to $R_X$.
In Section \@ref(package) we provide a basic description of how to quantify and potentially control the approximation error.
Further, the RNA-seq example in Section \@ref(examples) provides an illustration of this in practice.
